---
title: "Interential Statistics Project - Part 2"
author: "Mauricio Noris Freire"
date: "November 21, 2015"
output: pdf_document
---

## Loading the ToothGrowth database

```{r, echo=TRUE}

data(ToothGrowth)
str(ToothGrowth)


```

### Documentation 
According with the documentation, the dataset correspondes of the response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid). 

#### Format

A data frame with 60 observations on 3 variables. 

[,1]  len  numeric  Tooth length 
[,2]  supp  factor  Supplement type (VC or OJ). 
[,3]  dose  numeric  Dose in milligrams.  


## Provide a basic summary of the data.


```{r, echo=TRUE}

summary(ToothGrowth)

```


## Confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 


```{r, echo=TRUE}

boxplot(  len~supp*dose
        , data=ToothGrowth
        , notch=FALSE
        , col=(c("gold","darkgreen"))
        , main="Tooth Growth"
        , xlab="Suppliment and Dose")

```

## State your conclusions and the assumptions needed for your conclusions. 

There is a correlation between the dosage and the tooth size and the orange juice is more effective than ascorbic acid for small dosages and almost the same effect for the biggest dosage.
