---
title: "Inferential Statistics Project"
author: "Mauricio Noris Freire"
date: "November 19, 2015"
output: pdf_document
---

## Overview: 
Investigation the exponential distribution in R compared with CLT
by simulation of 1000 randomically generated numbers. Comparison of theorical versus obtained from the sample.


## Simulations 
Include English explanations of the simulations you ran, with the accompanying R code. Your explanations should make clear what the R code accomplishes. Using rexp(n, lambda) where lambda is = 0.2  as the rate parameter, and the variance is obtained by 1/lambda. 

### Setting the simulations parameters

```{r, echo=TRUE}
params.seed   <- 555555
params.n_sim  <- 1000
params.n_dist <- 40
params.lambda <- 0.2
params.precision <- 2

set.seed(params.seed)

simulate <- function(f) {
  r <- NULL
  for (s in 1 : params.n_sim) { 
    r = c(r, f(rexp(params.n_dist, params.lambda)))   
  }
  r
}
```

## Sample Mean versus Theoretical Mean
Used 40 random exponential distribution for 1000 collections . Also, the sample and theoretical means are reported for the comparison.

```{r}
output.means <- simulate(mean)
output.means.sample <- round(mean(output.means), params.precision)
output.means.theorical <- 1/params.lambda

hist(output.means, main="Dist 1000 averages of 40 exp"
                 , xlab= "Mean"
                 , ylab="Density"
                 , breaks= 30
                 , col="gray")
abline(v= output.means.sample, col="red",lwd= 2)

matrix(data=c(output.means.theorical, output.means.sample)
       , nrow=1
       , ncol=2
       , byrow=TRUE
       , dimnames=list(c("Mean"), c("Theoretical", "Sample")))

```

## Sample Variance versus Theoretical Variance
Used 40 means of variances obtained from random exponential distribution for 1000 collections . Also, the sample and theoretical variances are reported for the comparison.

```{r}
output.vars  <- simulate(var)
output.vars.sample <- round(mean(output.vars), params.precision)
output.vars.theorical <- (1/params.lambda)^2
hist(output.vars, main="Dist 1000 vars of 40 exp"
     , xlab= "Var"
     , ylab="Density"
     , breaks= 30
     , col="gray")
abline(v= output.vars.sample, col="red",lwd= 2)

matrix(data=c(output.vars.theorical, output.vars.sample)
       , nrow=1
       , ncol=2
       , byrow=TRUE, dimnames=list(c("Var"), c("Theoretical", "Sample")))

```


## Distribution
To demonstsrate CLT we execute two more simulations and compared. First only random exponentials and then the mean of groups of 40 from a sample of 1000.

```{r, echo=TRUE}
output.exponentials <- rexp(params.n_sim, params.lambda)
output.exponentials.mean <- (mean(output.exponentials))

par(mfrow=c(1,2))
 
hist(output.exponentials, main="Dist of 1000 exp"
                        , xlab= "Random"
                        , ylab="Density"
                        , breaks= 50
                        , col="gray")
abline(v= output.exponentials.mean, col="red",lwd= 2)

h1<- hist(output.means  , breaks= 50
                        ,col="grey"
                        , main=""
                        , xlab= "Dist of 1000 grouped for 40 exp"
                        , ylab="Density")

  abline(v=mean(output.means),col="red",lwd=2)
    

          x_norm <-seq( min(output.means)
                      , max(output.means) , length= 100)
          
          y_norm<-dnorm( x_norm
                       , mean=mean(output.means)
                       , sd=sd(output.means))

          y_norm<-y_norm*diff(h1$mids[1:2])*length(output.means)

          lines(x_norm, y_norm,col="red",lwd=2)

```
